### 2024-05-10 paperinfo.tex

% Please do not leave any empty line before the closing braces (}).

\PaperTitle{An Integrated Approach for Wildlife Recognition in Nepal Using Video Along With Audio}

\Authors{
    Zishan Siddique \textsuperscript{a}, 
    Sandip Pariyar \textsuperscript{b}, 
    Venus Bastola\textsuperscript{c},
    Sushant Phagu \textsuperscript{d},
    Binay Lal Shrestha \textsuperscript{e},
    Pravin Sangroula \textsuperscript{f}
}

\Abstract{
Wildlife around the world is in decline primarily due to the loss of habitat as well as the intersection of territory between humans and wild animals. Manual recognition of animals
can be more accurate but will require exponentially greater resources both in terms of
capital and labor making it unfeasible in large-scale deployment, especially for a country like Nepal. We have developed a system for the recognition and classification of wild animals using deep convolutional neural network (DCNN)
architecture to aid conservation as well as as study of our ecological system. We used iNaturalist as our source for image data and Xeno-canto.org as the source for audio data. We were able to achieve an F1-Score of 86.12\% on image data of 44 species of animals and an F1-Score of 88.1\% on audio data of 23 species of birds all found in Nepal.
}

% If you don't have any keyword simply remove the text between the curly braces
% But DO NOT remove the \Keywords command
\Keywords{Convolutional Neural Network,EfficientNet, Audio, Video}

\Affiliation{
    \textsuperscript{a,b,c,d,e,f}
    \textit{Purwanchal Campus,IOE,Tribhuvan University,Nepal}
}
%\Affiliation{
  %  \textsuperscript{b}
  %  \textit{Affiliation of Author-2}

%\Affiliation{
   % \textsuperscript{c}
   % \textit{Affiliation of Author-3}
%}

\Email{
    \textsuperscript{a} 076bct095@ioepc.edu.np , 
    \textsuperscript{b} 076bct074@ioepc.edu.np ,
    \textsuperscript{c} 076bct091@ioepc.edu.np,
    \textsuperscript{d} 076bct090@ioepc.edu.np,
    \textsuperscript{d} binay@ioepc.edu.np,
    \textsuperscript{d} pravin@ioepc.edu.np

}